---
title: "Rejection Sampling"
author: "Jason Bub"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Samplr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
library(samplr)
library(ggplot2)
```

This package implements one and two dimensional rejection sampling for continuous probability distribution functions.
 
## One-dimensional

The easiest form of rejection sampling occurs using one-dimensional pdfs. For example, we can sample from a beta distribution, in this case Beta(2,1). We first define a function for this pdf,

```{r}
betapdf <- function(x) {
  ifelse(0 < x & x < 1, 2*x, 0)
}
```

We can then use the `samplr` function to sample from this distribution, 

```{r}
samplr(1,betapdf)
```

With many samples, we are able estimate the distribution by using `hist` to plot the density of the samples,

```{r}
hist(samplr(10000,betapdf), main = NULL)
```

We can also sample from distributions with unbounded support,

```{r}
f <- function(x) {
  if (x < 0) {
    1/2*exp(x)
  }
  else {
    1/2*exp(-x)
  }
}
hist(samplr(10000,f), main = NULL)
```

We can find probabilities associated with these pdfs using the `psamplr` function, which is similar to the built in R functions `pnorm`, `pbinom`, etc., which gives $P(X < q)$, for some $q$ specified in the calling of the  `psamplr` function.

```{r}
psamplr(betapdf,.5)
psamplr(f,0)
```

Expectation can also be calculated using the `expsamplr` function, which calculates $E[X]$.

```{r}
expsamplr(betapdf)
expsamplr(f)
```

## Two-dimensional

We can also use the `samplr` function to sample from two-dimensional pdfs as well, specifying the parameter `twod = TRUE`. In this case, the input of the function that you wish to sample from must be a vector containing the random variables `x` and `y`, for example,

``` {r}
g <- function(z) {
  x <- z[1]
  y <- z[2]
  ifelse(0 < x & x < 1 & 0 < y & y < 1, x + y, 0)
}
samplr(1, g, twod = TRUE)
```

We can get an idea of what this distribution looks like by finding many samples, and using `ggplot` with `geom_density_2d()` to get contours of the pdf.

``` {r}
samps <- data.frame(samplr(10000, g, twod = TRUE))
colnames(samps) <- c("x","y")
ggplot(samps, aes(x, y)) + geom_density_2d()
```

The `samplr` function also works for two-dimensional pdfs with unbounded support,

```{r}
h <- function(z) {
  x <- z[1]
  y <- z[2]
  ifelse(0 < x & 0 < y, exp(-x)*exp(-y), 0)
}
samps <- data.frame(samplr(10000, h, twod = TRUE))
colnames(samps) <- c("x","y")
ggplot(samps, aes(x, y)) + geom_density_2d()
```

Like for one-dimensional pdfs, probabilities can also be calculated for two-dimensional pdfs using the `psamplr` function with a second function used to take various combinations of the random variables `X` and `Y`. The second function, `g` in the parameters, must have an input that is a vector of the random variables. For two-dimensional pdfs, `psamplr` calculates $P(g(X,Y) < q)$.

```{r}
prob1 <- function(z) {
  x <- z[1]
  y <- z[2]
  x + y
}
prob2 <- function(z) {
  x <- z[1]
  y <- z[2]
  x*y
}
psamplr(g,1,prob1)
psamplr(g,1,prob2)
```

Expectation values can also be found for two-dimensional pdfs using the `expsamplr` with a second input function of a combination of the random varialbes, similar to the `psamplr` function. For two-dimensional pdfs, `expsamplr` calculates $E[g(X,Y)]$.

```{r}
expect <- function(z) {
  x <- z[1]
  y <- z[2]
  x
}
expsamplr(g,prob1)
expsamplr(g,expect)
```






